---
title: "Replicating Crooker, Azevedo and Fenn, 2020"
author: "John R. Crooker, Ph.D."
date: "June 29, 2020"
output: html_document
---

## Some functions 

This function will be called below.  It simply calculates summary statistics for numerical columns in a data.frame.

```{r}
my_summary_stats_datframe<-function(x) {
  my_matrix<-matrix(0,nrow=ncol(x),ncol=9)
  for (i in 1:ncol(x)) {
    i_smy<-summary(x[,i])
    my_matrix[i,1:length(i_smy)]<-as.numeric(i_smy)
    my_matrix[i,8]<-sum(!is.na(x[,i]))
    my_matrix[i,9]<-sd(x[,i],na.rm=TRUE)
  }
  my_smy<-cbind(cbind(my_matrix[,c(8,7)],my_matrix[,1:6]),my_matrix[,9])
  my_datfr<-data.frame(my_smy)
  names(my_datfr)<-c("Obs","Miss.","Min","1Q","Med.","Mean","3Q","Max","SD")
  rownames(my_datfr)<-names(x)
  return(my_datfr)
}

```

## Loading the data

The data is available in a [link](https://github.com/jcrooker/MLB-Sign-Stealing) Github repository. 

```{r}
model_data2<-read.csv("https://raw.githubusercontent.com/jcrooker/MLB-Sign-Stealing/master/mlb_bavg_data.csv",
                      header=TRUE,
                      stringsAsFactors = FALSE)

numeric_cols<-unlist(lapply(as.list(1:ncol(model_data2)),
                            function(i) {is.numeric(model_data2[,i])}))
smy_stats2_datfr<-my_summary_stats_datframe(x=model_data2[,numeric_cols])
print(smy_stats2_datfr)
```

Estimate the model

\[
\begin{split}
\text{(Eq 18)  }&bavg_{i,t}=\beta_0+\beta_1bavg_{i,t-1}+\beta_2age_{i,t}+\sum_{j=1}^{J}\theta_jI_{i,t,j}+\epsilon_{i,t},
\end{split}
\]

```{r}
mlb_bavg<-lm(bavg~Lbavg+age+Y11+Y12+Y13+Y14+Y15+Y16+Y17+Y18,
             data=model_data2)
coef_table<-summary(mlb_bavg)$coefficients

sigma<-round(summary(mlb_bavg)$sigma,4)
rsq<-round(summary(mlb_bavg)$r.squared,4)
adjrsq<-round(summary(mlb_bavg)$adj.r.squared,4)

print(summary(mlb_bavg))
```

Note that Heteroskedasticity is a problem and should be corrected.

```{r}
### White's Test for heteroskedasticity
### There is evidence of Heteroskedasticity

uhat2<-mlb_bavg$residuals^2
yhat<-predict(mlb_bavg)
yhat2<-yhat^2

whites_data<-data.frame(uhat2,yhat,yhat2)
aux_model<-lm(uhat2~yhat+yhat2,data=whites_data)
print(summary(aux_model))
```

A plot of the model residuals:

```{r,echo=FALSE}
hist(mlb_bavg$residuals,
     xlab="Model residuals",
     ylab="Probability",
     probability = TRUE,
     lwd=2,
     las=1,
     main="Model Residuals Histogram")

x_vals<-seq(min(mlb_bavg$residuals),max(mlb_bavg$residuals),length=1000)
y_val<-dnorm(x_vals,mean=0,sd=sigma)
points(x=x_vals,
       y=y_val,
       type="l",
       col="blue",
       lwd=2)
```

## The MLB Empirical Distribution

```{r}
difference_between_discrete_empirical_densities<-function(y1,y2) {
  n1<-sum(y1)
  n2<-sum(y2)
  E2<-(y1/n1)*n2
  chi_stat<-((y2-E2)^2)/E2
  return(sum(chi_stat))
}
```

The *Monte Carlo* algorithm:

```{r}
my_seasons<-unique(model_data2$year)
model_data2$team[model_data2$team=="FLA"]<-"MIA"

my_teams<-unique(model_data2$team)

my_matrix<-matrix(NA,nrow=length(my_seasons),
                  ncol=length(my_teams))
chi_critical<-qchisq(p=0.95,df=1,lower.tail=TRUE)

for (season in my_seasons) {
  for (team in my_teams) {
    st_dat<-model_data2[model_data2$year==season & model_data2$team==team,]
    st_dat$pred<-predict(mlb_bavg,newdata=st_dat)
    lineup_chi<-rep(NA,nrow(st_dat))
    for (j in 1:nrow(st_dat)) {
      j_hits<-st_dat$H[j]
      j_not<-st_dat$AB[j]-j_hits
      j_y1<-c(st_dat$pred[j],1-st_dat$pred[j])
      j_y2<-c(j_hits,j_not)
      lineup_chi[j]<-difference_between_discrete_empirical_densities(y1=j_y1,
                            y2=j_y2)
    }
    my_matrix[which(season==my_seasons),
              which(team==my_teams)]<-sum(lineup_chi>chi_critical,
                                          na.rm=TRUE)
  }
}

mlb_datfr<-data.frame(Seasons=my_seasons,
                      my_matrix)
names(mlb_datfr)[2:ncol(mlb_datfr)]<-my_teams
```

The empirical distribution of our test-statistic appears in the following histogram.

```{r,echo=FALSE,results='asis'}
hist(unlist(mlb_datfr[,-1]),probability=TRUE,
     xlab="Empirical Test-Statistic",
     main="Empirical Distribution for MLB, 2010-2018",
     las=1)
```

The frequency distribution of our the empirical statistic in tabular form.

```{r}
mlb_stat<-unlist(mlb_datfr[,-1])
omlb_stat<-mlb_stat[order(mlb_stat)]

mlb_95<-omlb_stat[floor(0.95*length(omlb_stat))]

stat_tab<-table(mlb_stat)
stat_datfr<-data.frame(Statistic=names(stat_tab),
                       Count=as.numeric(stat_tab),
                       Pct=as.numeric(stat_tab)/sum(as.numeric(stat_tab)))
print(stat_datfr)
```

